Update and evaluation of the ozone dry deposition in the OsloCTM3 v1.0

### 2018-12-10 DryDep.tex
Added more references and reorganized some text.

In Fig.~\ref{fig:mmm_drydep_season_pft}, the seasonal cycles of dry deposition velocities are shown for the nine surface categories. The patterns and absolute numbers differ substantially between the Wesely scheme and the EMEP scheme and the multi-model-mean. The divergence of the average dry deposition velocities between the Wesely scheme and the EMEP scheme in desert regions ($\Delta \overline{v}_\text{desert}^\chem{O_3} = 0.18\,\unit{cm\,s^{-1}}$) as well as grassland ($\Delta \overline{v}_\text{grassland}^\chem{O_3} = 0.54\,\unit{cm\,s^{-1}}$) is quite remarkable as well as the difference of both to the multi-model-mean in tropical forest regions ($\Delta \overline{v}_\text{tropical forest}^\chem{O_3} = 0.49\,\unit{cm\,s^{-1}}$). In contrast to the Wesely scheme and the EMEP scheme, the multi-model-mean indicates a rather pronounced seasonal cycle in desert regions ($0.10\,\unit{cm\,s^{-1}} \leq v^\chem{O_3}_\text{desert}\leq 0.15\,\unit{cm\,s^{-1}}$). The estimated dry deposition velocities over desert regions are consistent with the previously mentioned baseline values of ozone dry deposition, which means that in the EMEP scheme they are about one order of magnitude lower than in the Wesely scheme. From measurements of ozone fluxes in the Sahara desert, \citet{AE:Gusten1995} deduced $v^\chem{O_3}_\text{desert, day} = 0.1\,\unit{cm\,s^{-1}}$, $v^\chem{O_3}_\text{desert, night} = 0.04\,\unit{cm\,s^{-1}}$, and $\overline{v}^\chem{O_3}_\text{desert} = 0.065\,\unit{cm\,s^{-1}}$. This implies, that ozone dry deposition over desert regions is highly overestimated in the Wesely scheme as well as in TF~HTAP models, while it may be underestimated in the EMEP scheme. Similarly, the dry deposition velocities over water differ. From measurements over the ocean a mean value of $\overline{v}^\chem{O_3}_\text{water} = 0.019\,\unit{cm\,s^{-1}}$ has been deduced \citep{JGR:Helmig2012}. In a model study of different mechanisms of dry deposition to ocean waters by means of prescribed $\overline{v}_\text{water}^\chem{O_3}$ and one- and two-layer gas exchange modeling, \citet{ACP:Luhar2017} found $\overline{v}_\text{water}^\chem{O_3}$ ranging between $0.018\,\unit{cm\,s^{-1}}$ (two-layer scheme) and $0.039\,\unit{cm\,s^{-1}}$ (prescribed). With $\overline{v}^\chem{O_3}_\text{water} = (0.046\pm0.002)\,\unit{cm\,s^{-1}}$, the EMEP scheme (Section~\ref{subsubsec:Rb}) yields probably a too strong dry deposition to ocean which implies that ozone concentrations might even become larger and dry deposition even lower in the model if a two-layer scheme would be implemented. With respect to vegetation, we might be able to improve the model performance further by allowing more PFTs, especially in regions covered by tropical or coniferous forest. 

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Comparison with observations}
\label{subsec:obs}
In this section, we compare our results to observations. For all comparisons, we use the original resolution of the OsloCTM3 ($2.25^\circ\times 2.25^\circ$) instead of the re-gridded ($3^\circ\times 3^\circ$). Although we do not assess the impacts of differing resolution in models on compatibility with observation, we can assume that both, high and low biases, may occur.

In Fig.~\ref{fig:mmm_drydep_stations}a), seasonal cycles of average ozone dry deposition fluxes for six observation sites are shown. We have computed a model average for all sensitivity studies at the closest grid point and show the $1\,\sigma$ error band. The shaded area around the multi-model-mean indicates the broad range of model results but is not an actual error band since such is not given in \citet{ACP:Hardacre2015}. At about $4$ of $6$ sites, the EMEP scheme performs better than the Wesely scheme and better than or the same as the multi-model-mean. We use a $\chi^2$-test
\begin{equation}
  \chi^2 = \sum_{i=1}^{12}\frac{\left(\overline{\chem{O_3}}_{\text{DD,\,}i}^\text{sim}-\overline{\chem{O_3}}_{\text{DD,\,}i}^\text{obs}\right)^2}{\sigma_i^2},
\end{equation}
with an estimated standard deviation of observation $\sigma_i=1\,\unit{mmol\,m^{-2}\,s^{-1}}$ and divide it by the number degrees of freedom (NDF) to assess this subjective analysis in a more objective way. The closer to $1$ this test scores, the better does the simulation represent the observation. A score between $0$ and $1$ indicates that the estimated $\sigma$ is too small. The results of the $\chi^2$-test are shown together with the divergences in Fig.~\ref{fig:mmm_drydep_stations}b). The $\chi^2$-test reveals also that in $4$ of $6$ cases the EMEP scheme improves the performance of the OsloCTM3 with respect to ozone dry deposition fluxes, although a satisfying result is only achieved for two sites (Castel Porziano, Blodgett Forest).
With only one full year of simulation, the model uncertainty regarding the seasonal cycle at observational sites cannot be quantitatively assessed. Further more, the observational averages comprise at most nine years worth of data. Statistically, these data may still subject to interannual variability. Among other aspects, the horizontal as well as vertical resolution play an important role in the model performance. In the rather coarse resolution, good matches between observation and model are only to be expected if the station is at a location which is representative for an area similar to the model resolution.

@@

Future work on the OsloCTM3 should include a direct output of dry deposition velocities for diagnostic purposes. The model performance could also be improved by allowing for more plant functional types than currently used or implementing an actual photosynthesis-based modeling of plants (Deposition of Ozone for Stomatal Exchange -- DO3SE). The photolysis- and chemical reaction computation as well as reaction rates shall also be revised in the future.

>>>>

In Fig.~\ref{fig:mmm_drydep_season_pft}, the seasonal cycles of dry deposition velocities are shown for the nine surface categories. The patterns and absolute numbers differ substantially between the Wesely scheme and the EMEP scheme and the multi-model-mean. The divergence of the average dry deposition velocities between the Wesely scheme and the EMEP scheme in desert regions ($\Delta \overline{v}_\text{desert}^\chem{O_3} = 0.18\,\unit{cm\,s^{-1}}$) as well as grassland ($\Delta \overline{v}_\text{grassland}^\chem{O_3} = 0.54\,\unit{cm\,s^{-1}}$) is quite remarkable as well as the difference of both to the multi-model-mean in tropical forest regions ($\Delta \overline{v}_\text{tropical forest}^\chem{O_3} = 0.49\,\unit{cm\,s^{-1}}$). In contrast to the Wesely scheme and the EMEP scheme, the multi-model-mean indicates a rather pronounced seasonal cycle in desert regions ($0.10\,\unit{cm\,s^{-1}} \leq v^\chem{O_3}_\text{desert}\leq 0.15\,\unit{cm\,s^{-1}}$). The estimated dry deposition velocities over desert regions are consistent with the previously mentioned baseline values of ozone dry deposition, which means that in the EMEP scheme they are about one order of magnitude lower than in the Wesely scheme. From measurements of ozone fluxes in the Sahara desert, \citet{AE:Gusten1995} deduced $v^\chem{O_3}_\text{desert, day} = 0.1\,\unit{cm\,s^{-1}}$, $v^\chem{O_3}_\text{desert, night} = 0.04\,\unit{cm\,s^{-1}}$, and $\overline{v}^\chem{O_3}_\text{desert} = 0.065\,\unit{cm\,s^{-1}}$. This implies, that ozone dry deposition over desert regions is highly overestimated in the Wesely scheme as well as in TF~HTAP models, while it may be underestimated in the EMEP scheme. Similarly, the dry deposition velocities over water differ. From measurements over the ocean a mean value of $\overline{v}^\chem{O_3}_\text{water} = 0.019\,\unit{cm\,s^{-1}}$ has been deduced \citep{JGR:Helmig2012}. In a model study of different mechanisms of dry deposition to ocean waters by means of prescribed $\overline{v}_\text{water}^\chem{O_3}$ and one- and two-layer gas exchange modeling, \citet{ACP:Luhar2017} found $\overline{v}_\text{water}^\chem{O_3}$ ranging between $0.018\,\unit{cm\,s^{-1}}$ (two-layer scheme) and $0.039\,\unit{cm\,s^{-1}}$ (prescribed). With $\overline{v}^\chem{O_3}_\text{water} = (0.046\pm0.002)\,\unit{cm\,s^{-1}}$, the EMEP scheme (Section~\ref{subsubsec:Rb}) yields probably a too strong dry deposition to ocean which implies that ozone concentrations might even become larger and dry deposition even lower in the model if a two-layer scheme would be implemented. With respect to vegetation, we might be able to improve the model performance further by allowing more PFTs and phenologies, especially in regions covered by tropical or coniferous forest \citep{GCB:Anav2017}. 

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Comparison with observations}
\label{subsec:obs}
In this section, we compare our results to observations. For all comparisons, we use the original resolution of the OsloCTM3 ($2.25^\circ\times 2.25^\circ$) instead of the re-gridded ($3^\circ\times 3^\circ$).

In Fig.~\ref{fig:mmm_drydep_stations}a), seasonal cycles of average ozone dry deposition fluxes for six observation sites are shown. We have computed a model average for all sensitivity studies at the closest grid point and show the $1\,\sigma$ error band. The shaded area around the multi-model-mean indicates the broad range of model results but is not an actual error band since such is not given in \citet{ACP:Hardacre2015}. At about $4$ of $6$ sites, the EMEP scheme performs better than the Wesely scheme and better than or the same as the multi-model-mean. We use a $\chi^2$-test
\begin{equation}
  \chi^2 = \sum_{i=1}^{12}\frac{\left(\overline{\chem{O_3}}_{\text{DD,\,}i}^\text{sim}-\overline{\chem{O_3}}_{\text{DD,\,}i}^\text{obs}\right)^2}{\sigma_i^2},
\end{equation}
with an estimated standard deviation of observation $\sigma_i=1\,\unit{mmol\,m^{-2}\,s^{-1}}$ and divide it by the number degrees of freedom (NDF) to assess this subjective analysis in a more objective way. The closer to $1$ this test scores, the better does the simulation represent the observation. A score between $0$ and $1$ indicates that the estimated $\sigma$ is too small. The results of the $\chi^2$-test are shown together with the divergences in Fig.~\ref{fig:mmm_drydep_stations}b). The $\chi^2$-test reveals also that in $4$ of $6$ cases the EMEP scheme improves the performance of the OsloCTM3 with respect to ozone dry deposition fluxes, although a satisfying result is only achieved for two sites (Castel Porziano, Blodgett Forest).
With only one full year of simulation, the model uncertainty regarding the seasonal cycle at observational sites cannot be quantitatively assessed. Further more, the observational averages comprise at most nine years worth of data. Statistically, these data may still subject to interannual variability. Among other aspects, the horizontal as well as vertical resolution play an important role in the model performance. Although, we do not explicitly assess the impacts of differing resolutions in our model, we can assume that both, high and low biases, may exist depending on the location of the observation side and the distribution of local sources \citep{AE:Schaap2015}. Since we use a rather coarse resolution, good matches between observation and model are only to be expected if the station is at a location which is representative for an area similar to the respective model gridbox.

@@

Future work on the OsloCTM3 should include a direct output of dry deposition velocities for diagnostic purposes. The model performance could also be improved by allowing for more plant functional types and phenologies than currently used or implementing an actual photosynthesis-based modeling of plants (Deposition of Ozone for Stomatal Exchange -- DO3SE). The photolysis- and chemical reaction computation as well as reaction rates shall also be revised in the future.
